---
title: "Clustering based on businesses strategic actions"
runningheader: "Clustering tourism businesses based on HHS Measures Adopted" # only for pdf output
subtitle: "An implementation in hierarchical clustering technique" # only for html output
author: "Arogya Koirala"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: 
    tufte_features: ["fonts", "background", "italics"]
  tufte::tufte_handout:
    fig_width: 12
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
---
```{css, echo=FALSE}
.dataTables_wrapper {
    position: relative;
    width: 60%;
    clear: both;
    zoom: 1;
}
```

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=15, fig.height=6) 
```



# Why?
In our survey, we’ve collected information around the kind of health and sanitation practices in place at different tourism businesses in and around Kathmandu. We want to use simple clustering techniques to group businesses based on their health and sanitation preparedness, and use maps to spatially represent this information.
Since our respondents are only a small subset of tourism businesses in Kathmandu, we also want to overlay OSM data and see if any patterns emerge.


# Setup

Import necessary libraries.

```{r message=FALSE, results='hide'}
library(vegan)
library(ggplot2)
library(purrr)
library(rworldmap)
library(leaflet)
library(cluster)
library(UpSetR)
library(DT)
# Imports
source("~/projects/c2m2/kathmandu-survey/utils/functions.R")
source("~/projects/c2m2/kathmandu-survey/utils/constants.R")
```

# Data import 
In this step, we import the latest business survey dataset, select relavant columns for our analysis, and rename columns for better legibility.

```{r}

# Get survey data
survey_data_path <- paste0(
  ROOT_URL, 
  "raw/data/business_data_20210531_wloc.xlsx"
)

businesses <- IO.XlsSheetToDF(
  excel_sheets(survey_data_path)[1], 
  survey_data_path
)

# Create flag variable to distinguish
# permanently closed businesses
businesses <- businesses %>% 
  mutate(d_shut_down_perm = ifelse(
    (i_econ_stop_business == 3), T, F))

# Subset data to only include HHS variables
hhs_vars <- c("X_index",
  "d_shut_down_perm",
  "o_expectd_problms_next_6_mnths__1",
  "o_expectd_problms_next_6_mnths__2",
  "o_expectd_problms_next_6_mnths__3",
  "o_expectd_problms_next_6_mnths__4",
  "o_expectd_problms_next_6_mnths__5",
  "o_expectd_problms_next_6_mnths__6",
  "o_expectd_problms_next_6_mnths__7",
  "m_biz_type",
  "m_coodinates", 
  "X_m_coodinates_latitude",
  "X_m_coodinates_longitude",
  "m_name_business"
) 

# Rename columns
businesses_hhs <- businesses %>% select(all_of(hhs_vars))
colnames(businesses_hhs) <- c(
  "id",
  "perm_closed",
  "pay_loans",
  "get_funds",
  "pay_taxes",
  "op_cost",
  "labor",
  "get_custs",
  "none",
  "type",
  "coordinates",
  "X_m_coodinates_latitude",
  "X_m_coodinates_longitude",
  "name"
)

renamed_cols <- colnames(businesses_hhs)

businesses_hhs <- businesses_hhs %>% filter(perm_closed == F)

# Preview
knitr::kable(
  businesses_hhs[1:6, 1:6], caption = 'A subset of businesses_hhs'
)
```

# Exploratory data analysis using the UpsetR package

The package `UpsetR`, which provides intersection totals for a combination of binary variables can be used to quickly scan through data.

```{r, fig.fullwidth=TRUE}

# Subset columns of interest
subs <- businesses_hhs %>% select(c(colnames(businesses_hhs)[3:11]))

# Function to convert all factors to numeric
asNumeric <- function(x) as.numeric(as.character(x))
factorsNumeric <- function(d) modifyList(
  d, 
  lapply(d[, sapply(d, is.factor)],asNumeric)
)

# Visualize using UpsetR 
upset(factorsNumeric(subs), 
      order.by="freq", 
      main.bar.color = "#c3092b", 
      sets.bar.color = "#c3092b", 
      group.by = "degree"
)
```

# Performing hierarchical clustering

Link 1: Approach, https://uc-r.github.io/hc_clustering

Link 2: We are going to use Jaccard distances (see, https://towardsdatascience.com/user-similarity-with-binary-data-in-python-d15940a702fc)




```{r}

# Subset only relevant columns
measures_var <- renamed_cols[3:9]
df <- businesses_hhs %>% 
  select(all_of(measures_var))

# Convert columns to numeric values
indx <- sapply(df, is.factor)
df[indx] <- lapply(df[indx], function(x) as.numeric(as.character(x)))

# Compute distance matrix based on "jaccard" distance
dist.mat<-vegdist(
  df,
  method="jaccard",
  binary = T
) 


# methods to assess
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

# function to compute coefficient
ac <- function(x) {
  agnes(df, method = x)$ac
}

map_dbl(m, ac)
```


```{r, fig.fullwidth=TRUE}
# Perform and plot clustering results
clust.res<-hclust(dist.mat, method = "ward")
plot(clust.res, hang=-1, cex=0.8)
```

## Notes:

1. We will be using k=5 to assign clusters


# Assigning clusters and profiling

Our next step is to assign cluster values to the dataset, and then generate statistics for each cluster to get a sense of cluster characteristics. Before we do that, however, lets look at percentage values for different columns for the entire population.

```{r, width=6}

totals <- df %>% summarise(
  n = n(),
  pay_loans=round(sum(as.numeric(as.character(pay_loans)))/n(), 2),
  get_funds=round(sum(as.numeric(as.character(get_funds)))/n(),2),
  pay_taxes=round(sum(as.numeric(as.character(pay_taxes)))/n(),2),
  op_cost=round(sum(as.numeric(as.character(op_cost)))/n(),2),
  labor=round(sum(as.numeric(as.character(labor)))/n(),2),
  get_custs=round(sum(as.numeric(as.character(get_custs)))/n(), 2),
  none=round(sum(as.numeric(as.character(none)))/n(),2)
)

datatable(totals, options = list(
  autoWidth = F
))
```

Now then, lets map our businesses to their respective clusters and look at how these values are distributed at the cluster level.

```{r}

sub_grp <- cutree(clust.res, k = 5)
df <- cbind(df, sub_grp)
businesses_hhs <- cbind(businesses_hhs, sub_grp)

cluster_profiles <- df %>% 
  group_by(sub_grp) %>% 
  summarise(
    n = n(),
    pay_loans=sum(as.numeric(as.character(pay_loans)))/n(),
    get_funds=sum(as.numeric(as.character(get_funds)))/n(),
    pay_taxes=sum(as.numeric(as.character(pay_taxes)))/n(),
    op_cost=sum(as.numeric(as.character(op_cost)))/n(),
    labor=sum(as.numeric(as.character(labor)))/n(),
    get_custs=sum(as.numeric(as.character(get_custs)))/n(),
    none=sum(as.numeric(as.character(none)))/n()
)

datatable(cluster_profiles, options = list(
  autoWidth = F
))
```

It also makes sense to look at the different business types allocated to each cluster.

```{r, fig.fullwidth=TRUE,  fig.height=7}
ggplot(
  businesses_hhs, 
  aes(x=factor(type))
  ) + 
  geom_bar(
  stat="count", 
  width=0.7, 
  fill="#c3092b") +
  facet_grid(sub_grp ~ .) + 
  scale_x_discrete(
  breaks=1:9,
  labels=c(
    "Hotel",
    "Rest & B",
    "Trav. Tour",
    "Handicraft",
    "Rafting",
    "Trekking",
    "Mountnring", 
    "Shop", 
    "Other"
    )
  ) +
  theme_linedraw()
```

# Viewing clusters on a map

```{r}

# Let's assign some names to these clusters
businesses_hhs <- businesses_hhs %>%
  mutate(
    cluster_name = ifelse(
      sub_grp == 1, "Worried abt everything",
      ifelse(
        sub_grp == 2, "Only worried about paying loans",
        ifelse(
          sub_grp == 3, "Worried about operating and getting customers",
          ifelse(
            sub_grp == 4, "Financially worried (loans, funds, taxes)",
            ifelse(
              sub_grp == 5, "Mixed/unclear",
              ifelse(
                sub_grp == 6, "Six", "Seven"
              )
            )
          )
        )
      )
    )
  )


# Cover lat/lng to numeric values and visualize
businesses_hhs$lat <-as.numeric(
  as.character(businesses_hhs$X_m_coodinates_latitude)
  )

businesses_hhs$lng <- as.numeric(
  as.character(    businesses_hhs$X_m_coodinates_longitude)
  )

onlyLatLng <- businesses_hhs[!is.na(businesses_hhs$lat), ]
onlyLatLng <- businesses_hhs[!is.na(businesses_hhs$lng), ]

colors <- c(
  "#e41a1cff", 
  "#377eb8ff", 
  "#4daf4aff", 
  "#984ea3ff",
  "#ff7f00ff", 
  "#ffff33ff", 
  "#a65628ff"
  )
factpal <- colorFactor(colors, onlyLatLng$cluster_name)

leaflet(data = onlyLatLng, width = "70%") %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    lng = ~lng, 
    lat = ~lat, 
    color = ~factpal(cluster_name),
    stroke = FALSE, 
    fillOpacity = 0.5,
    label = ~name) %>%
  addLegend(
    "bottomright", 
    pal = factpal, 
    values = ~cluster_name,
    title = "Cluster Assignment",
    opacity = 1
  )



```